---
title: "Tools"
tutorial:
  id: "tools"
output:
  learnr::tutorial:
      progressive: true
      allow_skip: true
runtime: shiny_prerendered
description: "Tools Tutorial"
---

```{r setup, include=FALSE}
library(tidyverse)
library(PPBDS.data)
library(learnr)
library(shiny)
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
options(tutorial.exercise.timelimit = 60, tutorial.storage="local")
```

1) Seems weird that the set up chunk only has to include learnr. Does it really work without at least library(shiny)?

2) We need some commands related to git. Mainly just, issue the git commands we told you to, then type "git config user.name" and paste what comes back. Then, type "git config user.email" and paste what comes back. When you paste, give us both the line on which you typed the command and the line(s) below it. Add some git_sitrep() questions, which requires telling them to install **usethis**.

3) Let's make this more like the other tutorials. There are two groups: The Terminal and Git. Each group has a collection of exercises.

4) Should we do more, something with Github? Something with a PDF file? That would certainly get students going to the first set of Study Halls!

## Welcome!

In this tutorial, we will be focusing on the terminal window. The terminal window allows you to interact with the filesystem within your computer. Remember, if you ever have questions regarding how to navigate terminal, visit the "Tools" section of the primer .

## Name
###

``` {r name}
question_text(
  "Student Name:",
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Modify your answer",
  incorrect = "Ok"
)
```

## Getting Started

Let's open up the Terminal tab on the left window right next to "Console". We need to check that git has been installed first. Copy/paste 'git --version' into the terminal and press return.

You should get an output of the git version (ex. git version 2.24.3). If you do not, go to the [Tools section](https://davidkane9.github.io/PPBDS/tools.html) of the primer and follow the directions for installing git.

## Terminal Commands
### Working Directory: pwd


Now that you have Git installed, let's get started. First, we need to figure out where we are. To see your current folder, type 'pwd' (previews working directory). Hit return, and copy/paste what you see into the space below.

```{r pwd, echo=FALSE}
question_text("copy/paste here:",
              answer(NULL, correct = TRUE),
              incorrect = "Ok",
              try_again_button = "Modify your answer",
              allow_retry = TRUE)
```


### Seeing Items in the Directory: ls


So now that we know what folder we are in...how can we see it? Type 'ls' in the terminal and hit return. You should see all the items that are located in the bottom right window under "Home". Copy/paste what you see into the space below.

```{r ls, echo=FALSE}
question_text("copy/paste here:",
              answer(NULL, correct = TRUE),
              incorrect = "Ok",
              try_again_button = "Modify your answer",
              allow_retry = TRUE)
```

### Changing Directories:cd
**Part I.**
To open/close folders we use 'cd'. For our purposes, we will be opening or "changing" into our Desktop directory. Type in the command 'cd Desktop/' and copy/paste what you see into the space below.
```{r cd, echo=FALSE}
question_text("copy/paste here:",
              answer(NULL, correct = TRUE),
              incorrect = "Ok",
              try_again_button = "Modify your answer",
              allow_retry = TRUE)

```

**Part II.**
Awesome. Now type 'ls'. What do you think will happen?

### Make/Remove Directories: mkdir, rmdir
Ok, time to get organized. Let's trying making a folder called Gov-50 using the command 'mkdir Gov-50' and hit return. Use "ls" to see your newly created folder!


### Creating files: touch
Now we are going to create a file. Type in 'touch: text.txt' and hit return. Now click ls. Copy/paste what you see in the space below.

```{r touch, echo=FALSE}
question_text("copy/paste here:",
              answer(NULL, correct = TRUE),
              incorrect = "Ok",
              try_again_button = "Modify your answer",
              allow_retry = TRUE)

```

### Moving Files: mv
Alright...lets try moving our text.txt file to our Gov-50 folder using 'mv test.txt Gov-50/'. Now let's check! Use 'cd' to open our Gov-50 folder and ls to show its contents. Copy/paste what you see below!

```{r mv, echo=FALSE}
question_text("copy/paste here:",
              answer(NULL, correct = TRUE),
              incorrect = "Ok",
              try_again_button = "Modify your answer",
              allow_retry = TRUE)
```

### Copying Files: cp

Now try copying the test.txt file into your Gov-50 folder. Use 'cp test.txt Gov-50/.


### Removing Files:rm

Now lets try removing that same text.txt file with 'rm'. Use 'ls' to make sure it is gone!

## Submit

Some of these exercises were taken from the collection of [RStudio Primers](https://rstudio.cloud/learn/primers), a great resource for practicing your skills.

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```


