---
title: "Visualization"
output:
  learnr::tutorial:
      progressive: true
      allow_skip: true
runtime: shiny_prerendered
description: "Chapter 1 tutorial"
---

<!-- Ch. 1 tutorial questions written by Jessica Edwards  -->
<!-- Also taken from rstudio/learnr and rstudio-education/primers GitHub repos -->

```{r setup, include=FALSE}
library(tidyverse)
library(nycflights13)
library(PPBDS.data)
library(learnr)
library(learnrhash)
library(shiny)
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

## What are R and RStudio?

```{r name, echo=FALSE}
question_text(
  "Student Name:",
  answer(NULL, correct = TRUE),
  incorrect = "Ok",
  try_again_button = "Modify your answer",
  allow_retry = TRUE
)
```

RStudio is an Integrated Development Environment for R. What does that mean? Well, if you think of R as a language, which it is, you can think of RStudio as a program that helps you write and work in the language. RStudio makes programming in R much easier and we suggest that you use it!

### Test your knowledge

```{r quiz3, echo=FALSE}
quiz(
  question("What are R and RStudio?",
             answer("R is a programming language and RStudio is an integrated development environment (IDE)", 
                    correct = TRUE),
             answer("RStudio is a programming language and R is an IDE"),
             answer("R and RStudio are both programming languages"),
             answer("R and RStudio are both IDEs"),
             allow_retry = TRUE
    ),
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use R.", correct = TRUE, message = "RStudio organizes your workspace and makes it easier to write, use, debug, and save R code. I highly recommend it and will rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code", message = "And thank goodness! Code provides a reproducible record of your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel."),
    answer("Another name for R", message = "R and RStudio are two separate things. R is a language, like English. Think of RStudio as a program that helps you use the language, kind of like how a word processing program helps you write in English."),
    allow_retry = TRUE
  ),
  question("Do you need to install R if you already have RStudio?",
    answer("Yes.", correct = TRUE),
    answer("No.", message = "R does not come with RStudio; you need to install R separately.")
  ),
  question("In what order do you install R and RStudio?",
             answer("First R then RStudio", correct = TRUE),
             answer("First RStudio then R"),
             allow_retry = TRUE
    )
)
```

## How Do I Code in R?

Data frames and tibbles are R's structures for storing tabular data; if you inherit a tabular dataset in R, it will almost certainly come as one of these structures. 

Here, you will learn how to do three things with data frames and tibbles:

1. Look at the contents of a data frame or tibble
2. Open a help page that describes a data frame or tibble
3. Identify the variables and their types in a tibble 

### What is a data frame?

A __data frame__ is a rectangular collection of values, usually organized so that variables appear in the columns and observations appear in rows. 

Here is an example: the `mpg` data frame contains observations collected by the US Environmental Protection Agency on 38 models of cars. To see the `mpg` data frame, type `mpg` in the code chunk below and then click "Submit Answer."

```{r mpg-setup}
mpg <- as.data.frame(mpg)
```

```{r mpg, exercise = TRUE}

```

<div id="mpg-hint">
**Hint:** Type `mpg` and then click the Run Code button.
</div>

###  

Did you notice how much information was inside `mpg`? Me too. Sometimes the contents of a data frame do not fit on a single screen, which makes them difficult to inspect. We'll look at an alternative to using and examining data frames soon. But first let's get some help...

### Help pages

You can learn more about `mpg` by opening its help page. The help page will explain where the `mpg`dataset comes from and what each variable in `mpg` describes. To open the help page, type `?mpg` in the code chunk below and then click "Run Code".

```{r help, exercise = TRUE}

```

<div id="help-hint">
**Hint:** Type `?mpg` and then click the Submit Answer button.
</div>

### ? syntax

You can open a help page for any object that comes with R or with an R package. To open the help page, type a `?` before the object's name and then run the command, as you did with `?mpg`. This technique works for functions, packages, and more. 

Notice that objects created by you or your colleagues will not have a help page (unless you make one).

### What is a tibble?

The `flights` data frame in the nycflights13 package is an example of a _tibble_. Tibbles are a data frames with some extra properties. 

To see what I mean, use the code chunk below to print the contents of `flights`.

```{r flights, exercise = TRUE}

```

<div id="flights-hint">
**Hint:** Type the name of the data frame that you want to print and then click the Run Code button. We've already loaded the nycflight13 package for you.
</div>

###  

Good Job. `flights` describes every flight that departed from New York City in 2013. The data comes from the [US Bureau of Transportation Statistics](http://www.transtats.bts.gov/DatabaseInfo.asp?DB_ID=120&Link=0), and is documented in `?flights`.


### The tibble display

You might notice that `flights` looks a little differently than `mpg`. `flights` shows only the first few rows of the data frame and only the columns that fit on one screen.

`flights` prints differently because it's a __tibble__. Tibbles are data frames that are slightly tweaked to be more user-friendly. For example, R doesn't try to show you all of a tibble at once (but it will try to show you all of a data frame that is not a tibble).

You can use `as_tibble()` to return a tibble version of any data frame. For example, this would return a tibble version of `mpg`: `as_tibble(mpg)`. 

### Data types
 
```{r flights3, echo = FALSE}
flights
``` 
 
Did you notice that a row of three (or four) letter abbreviations appears under the column names of `flights`? These abbreviations describe the _type_ of data that is stored in each column of `flights`:

* `int` stands for integers.

* `dbl` stands for doubles, or real numbers.

* `chr` stands for character vectors, or strings.

* `dttm` stands for date-times (a date + a time).

There are three other common types of variables that aren't used in this dataset but are used in other datasets:

* `lgl` stands for logical, vectors that contain only `TRUE` or `FALSE`.

* `fctr` stands for factors, which R uses to represent categorical variables
  with fixed possible values.

* `date` stands for dates.

This row of data types is unique to tibbles and is one of the ways that tibbles try to be more user-friendly than data frames.

### Test your knowledge

```{r summary, echo=FALSE}
quiz(
  question("Which types of variables does `flights` contain? Check all that apply.",
         type = "multiple",
         allow_retry = TRUE,
         incorrect = "Not quite right. Look a little closer at `flights`.",
         answer("integers", correct = TRUE),
         answer("doubles", correct = TRUE),
         answer("factors"),
         answer("characters", correct = TRUE),
         correct = "Great Job!"
), question("Which of the following is NOT the name of a data type in R?",
           answer("Logicals"),
           answer("Integers"),
           answer("Characters"),
           answer("Strings", correct = TRUE),
           allow_retry = TRUE
  ),
  question("What is the operator used in R to test for inclusion?",
           answer("'in'"),
           answer("'%in%'", correct = TRUE),
           answer("'include'"),
           answer("'IN'"),
           allow_retry = TRUE
  ),
  question("What is the symbol used to bring up a help file?",
           answer("*"),
           answer("~"),
           answer("!"),
           answer("?", correct = TRUE),
           allow_retry = TRUE
  ),
  question("True / False: Seeing red text in the console is always bad because it denotes an error.",
           answer("True"),
           answer("False", correct = TRUE, message = "R will show red text in the console pane for errors, warnings, and messages. The only situation in which the code will not run is if there is an error."),
           incorrect = "R will show red text in the console pane for errors, warnings, and messages. The only situation in which the code will not run is if there is an error.",
           allow_retry = TRUE
  )
)
```

## What are R packages?

### Test your knowledge

```{r names, echo = FALSE}
quiz(caption = "Quiz - Working with Packages",
  question("What command do you use to install packages?",
    answer("`library()`", message = "We will see what `library()` does later."),
    answer("`install.packages()`", correct = TRUE),
    answer("`install_packages()`"),
    answer("There is no command. You must visit [cran.r-project.org](http://cran.r-project.org) and download packages manually.", message = "R makes it easy to download packages. You connect to the internet and then run one of the commands above."),
    allow_retry = TRUE
  ),
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R"),
    answer("Every time you restart your computer"),
    answer("Only once. Afterwards, R can find it on your hard drive as needed.", correct = TRUE),
    answer("Never, as long as you are connected to the internet.", message = "This could be true if you are using R over a cloud service. However, if you are using R locally on your own computer, you will need to install each package that you use on your computer."),
    allow_retry = TRUE
  ),
  question("Which of the following is NOT true about R packages?",
           answer("R packages extend the functionality of R by providing additional functions, data, and documentation"),
           answer("If comparing R and R packages to a phone, R would be the apps you can download on the phone and R packages would be the phone", correct = TRUE),
           answer("R packages are written by a worldwide community of R users"),
           answer("R packages can be downloaded for free from the Internet"),
           allow_retry = TRUE
  )
)
```

*Load the `ggplot2` package using `library()`.*

```{r exercise-1, exercise = TRUE, exercise.lines = 3}

```

## Explore your first datasets

### `trains` dataset

*Load the `PPBDS.data` package to look at the `trains` dataset.*
```{r exercise-2, exercise = TRUE, exercise.lines = 5}

  
trains
```

*Use the `glimpse()` function to look at the `trains` dataset. We already loaded the `dplyr` and `PPBDS.data` packages.*
```{r exercise-4, exercise = TRUE, exercise.lines = 5}
library(dplyr)
library(PPBDS.data)


```

```{r quiz-4, echo = FALSE}
quiz(caption = "Quiz - `trains` dataset",
  question("What is the data type of the `liberal` variable?",
           answer("character"),
           answer("logical", correct = TRUE),
           answer("double"),
           answer("factor"),
           allow_retry = TRUE
  ),
  question("How many rows does the `train` dataset have?",
           answer("8"),
           answer("15"),
           answer("115", correct = TRUE),
           allow_retry = TRUE
  )
)
```

*Extract the `income` variable in the `trains` dataset using the `$` operator.*
```{r exercise-5, exercise = TRUE}

```

### `qscores` dataset

*Look at the `qscores` dataset using `glimpse()`. We have loaded the `PPBDS.data` and the `dplyr` packages.*

```{r qscores, exercise = TRUE, exercise.lines = 5}

```

```{r quiz-5, echo = FALSE}
quiz(
  question("Which of the following is NOT an identification variable? (choose 2)",
           answer("`course_name`"),
           answer("`department`", correct = TRUE),
           answer("`workload`", correct = TRUE),
           answer("`course_number`"),
           allow_retry = TRUE
  )
)
```

## Data Visualization using the ggplot2 package

### Components of a graphic

```{r quiz-6, echo = FALSE}
quiz(
  question("What are the three essential components of a graphic?",
           answer("the dataset containing the variables in question", correct = TRUE),
           answer("the geometric object we can observe in a plot", correct = TRUE),
           answer("axes labels on a plot"),
           answer("the aesthetic attributes", correct = TRUE),
           allow_retry = TRUE
  ),
  question("What are the two important arguments that we need to provide the `ggplot()` function?",
           answer("`data` and `mapping`", correct = TRUE),
           answer("`data` and `aesthetics`"),
           answer("`data` and `layers`"),
           allow_retry = TRUE
  )
)
```

### Components of a scatterplot

*Look at this scatterplot made using the `trains` dataset. Higher numbers mean more conservative attitudes toward immigration-related policies on a 3 to 15 scale. Individuals with a treatment value of "Treated" were exposed to two Spanish-speakers on their regular commute. "Control" individuals were not.*

```{r trains}
ggplot(data = trains, mapping = aes(x = att_start, y = att_end, color = treatment)) +
  geom_point() + facet_wrap(~party)
```

```{r quiz-7, echo = FALSE}
quiz(
  question("Which `data` variable is mapped to the `x`-position `aes`thetic of the points?",
           answer("`att_start`", correct = TRUE),
           answer("`att_end`"),
           answer("`treatment`"),
           allow_retry = TRUE
  ),
  question("Which `data` variable is mapped to the `y`-position `aes`thetic of the points?",
           answer("`att_start`"),
           answer("`att_end`", correct = TRUE),
           answer("`treatment`"),
           allow_retry = TRUE
  ),
  question("Which `data` variable is mapped to the `color` `aes`thetic of the points?",
           answer("`att_start`"),
           answer("`att_end`"),
           answer("`treatment`", correct = TRUE),
           allow_retry = TRUE
  )
)
```

## Summary of Geoms

### Test Your Knowledge

```{r geoms, echo = FALSE}
quiz(
  question("Which plot(s) show the relationship between 2 numerical variables?",
           answer("scatterplot", correct = TRUE),
           answer("lineplot", correct = TRUE),
           answer("histogram"),
           answer("boxplot"),
           answer("barplot"),
           allow_retry = TRUE
  ),
  question("Which plot(s) show the distribution of 1 numerical variable?",
           answer("scatterplot"),
           answer("lineplot"),
           answer("histogram", correct = TRUE),
           answer("boxplot", correct = TRUE),
           answer("barplot"),
           allow_retry = TRUE
  ),
  question("Which plot(s) show the distribution of 1 categorical variable?",
           answer("scatterplot"),
           answer("lineplot"),
           answer("histogram"),
           answer("boxplot"),
           answer("barplot", correct = TRUE),
           allow_retry = TRUE
  ),
  question("Which of the following is NOT a geometric object?",
           answer("`geom_scatter()`", correct = TRUE),
           answer("`geom_line()`"),
           answer("`geom_histogram()`"),
           answer("`geom_boxplot()`"),
           answer("`geom_bar()`"),
           allow_retry = TRUE
  )
)
```

## Scatterplots

### Scatterplot with the `qscores` dataset

*Let's make a scatterplot using the `qscores` dataset. Load the `PPBDS.data` package and look at the `qscores` dataset by simply typing the name of the dataset.*

```{r exercise-6, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-6-hint}
library(...)

...
```

###

*Load the `ggplot2` package. Use the `ggplot()` function to create a scatterplot using the `qscores` dataset, `rating` on the x-axis and `hours` on the y-axis. Is there a relationship between the rating of a Harvard course versus the average hours per week of workload?*

```{r exercise-7, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-7-hint}
library(...)

ggplot(data = qscores, mapping = aes(x = ..., y = ...)) +
  ...
```

###

*Now we want to add a `size` `aes`thetic based on the number of students enrolled in each course. Add `size = enrollment` to the other `aes`thetics of the points. Do larger classes have a higher average overall course rating? What about higher workload?*

```{r exercise-8, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = rating, y = hours)) +
  geom_point()
```

```{r exercise-8-hint}
ggplot(data = qscores, mapping = aes(x = rating, y = hours, size = ...)) +
  geom_point()
```

### Overplotting

While this plot created using the `mpg` dataset only displays 126 points, it visualizes a data set that contains 234 points.

```{r echo = FALSE, out.width = "80%"}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

The missing points are hidden behind other points, a phenomenon known as _overplotting_. Overplotting is a problem because it provides an incomplete picture of the dataset. You cannot determine where the *mass* of the points fall, which makes it difficult to spot relationships in the data. 

### Causes of overplotting

Overplotting usually occurs for two different reasons:

1. The data points have been rounded to a "grid" of common values, as in the plot above
2. The dataset is so large that it cannot be plotted without points overlapping each other

How you deal with overplotting will depend on the cause. 

### Changing the transparency

If your overplotting is due to rounding, you can obtain a better picture of the data by making each point semi-transparent. For example you could _set_ the alpha aesthetic of the plot below to a _value_ less than one, which will make the points transparent. 

Try this now. Set the points to an alpha of 0.2, which will make each point 20% opague (i.e. five points staked on top of each other will create a solid black).

*Change the transparency of the points by setting `alpha = 0.2` within `geom_point()`.*

```{r exercise-9, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = rating, y = hours, size = enrollment)) +
  geom_point()
```

```{r exercise-9-hint}
ggplot(data = qscores, mapping = aes(x = rating, y = hours, size = enrollment)) +
  geom_point(alpha = ...)
```

### Adjusting the position

A second strategy for dealing with overplotting due to rounding is to adjust the position of each point. `geom_jitter()` plots a scatterplot and then adds a small amount of random noise to each point in the plot. It is a shortcut for adding a "jitter" position adjustment to a points plot (i.e, `geom_point(position = "jitter")`). Since the noise is random, it is unlikely that two points rounded to the same location will also be jittered to the same location. The result is a jittered plot that displays more of the data and provides a simple way to inspect patterns that occur in heavily gridded or overlapping data. 

*Replace `geom_point()` with `geom_jitter()` to create a jittered scatterplot.*

```{r exercise-10, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = rating, y = hours)) +
  geom_point()
```

### Strategies for large data

Alpha and jittering are less useful for large data. Jittering will not separate the points, and a mass of transparent points can still look black.

A better way to deal with overplotting due to large data is to visualize a summary of the data, by using a histogram or a linegraph, for example.

## Linegraphs

```{r linegraphs-setup, include = FALSE}
cces_data <- cces %>%
  filter(approval_ch %in% c("Approve / Somewhat Approve", "Disapprove / Somewhat Disapprove")) %>%
  group_by(year, approval_ch) %>%
  summarize(avg_age = mean(age))
```

*The `cces` dataset in the `PPBDS.data` package has been changed to show the average age of voters per year based on their approval of the president that year. The `cces` dataset was manipulated using the `filter()`, `group_by`, and `summarize()` functions within the `tidyverse` package, which you will learn how to use in the next chapter.*

```{r gov_qscores, echo = TRUE}
cces_data
```

*Use the `ggplot()` function to create a linegraph with the dataset called `cces_data` with `year` on the x-axis, `avg_age` on the y-axis, and `approval_ch` as the color aesthetic.*

```{r exercise-11, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-11-hint}
ggplot(data = cces_data, mapping = aes(x = ..., y = ..., color = ...)) +
  geom_line()
```

### Test your knowledge

```{r quiz-8, echo = FALSE}
quiz(
  question("True / False: Linegraphs and scatterplots both display the relationship between two numerical variables.",
           answer("True", correct = TRUE),
           answer("False"),
           allow_retry = TRUE),
  question("True / False: It is preferred to use scatterplots over linegraphs when the variable on the x-axis has an inherent ordering, like in a time series plot.",
           answer("True"),
           answer("False", correct = TRUE),
           allow_retry = TRUE)
)
```

## Histograms

### Test your knowledge

```{r quiz-9, echo = FALSE}
quiz(
  question("A histogram is:",
           answer("A plot that displays the relationship between two numerical variables"),
           answer("A plot that visualizes the distribution of a numerical variable", correct = TRUE),
           answer("A plot that visualizes the distribution of a categorical variable"),
           allow_retry = TRUE)
)
```

###

*Using the `qscores` dataset and `ggplot()`, make a histogram where `rating` is on the x-axis to see the distribution of Harvard course ratings.*

```{r exercise-12, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-12-hint}
ggplot(data = ..., mapping = aes(x = ...)) + ...
```

```{r quiz-10, echo = FALSE}
quiz(
  question_text("What is the default value of bins for this histogram?",
                answer("30", correct = TRUE), 
                allow_retry = TRUE)
)
```

###

*Add white vertical borders demarcating the bins by adding `color = "powderblue"` and `fill = "violet"` arguments to `geom_histogram()`.*

```{r exercise-13, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram()
```

```{r exercise-13-hint}
ggplot(data = qscores, mapping = aes(x = overall)) + 
  geom_histogram(color = ..., fill = ...)
```

###

*Specify the number of bins to be 10 via the `bins` argument and the `color` as "white" in `geom_histogram()`.*

```{r exercise-14, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram()
```

```{r exercise-14-solution}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram(bins = ..., color = ...)
```

###

*Specify the width of the bins to be 1 via the `binwidth` argument and the `color` as "white" in `geom_histogram()`.*

```{r exercise-15, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram()
```

```{r exercise-15-solution}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram(binwidth = 1, color = "white")
```

## Boxplots

```{r quiz-11, echo = FALSE}
quiz(
  question("Which of the following are included in the five-number summary of a numerical variable used to construct a boxplot?",
           answer("Minimum", correct = TRUE),
           answer("First quartile", correct = TRUE),
           answer("Median", correct = TRUE),
           answer("Mean"),
           answer("Third quartile", correct = TRUE),
           answer("Maximum", correct = TRUE)
  )
)
```

*Let's keep playing with the `qscores` dataset in the `PPBDS.data` package. We have filtered the dataset using the `tidyverse` package based on the five departments with the classes with the most enrollment: ECON, COMPSCI, STAT, GOV, and MATH. (You will learn how to do this in the next chapter). The filtered dataset is called `top_five_qscores`.*

```{r boxplot-setup, include = FALSE}
top_five_enrolled <- qscores %>% 
  group_by(department) %>%
  summarize(enrollment = sum(enrollment)) %>% 
  arrange(desc(enrollment)) %>%
  pull(department)

top_five_qscores <- qscores %>% filter(department %in% top_five_enrolled[0:5])
```

```{r boxplot-top-five-qscores, echo = TRUE}
top_five_qscores
```

*Using the `top_five_qscores` dataset and `ggplot()`, make a boxplot where `department` is on the x-axis and `overall` is on the y-axis.*

```{r exercise-16, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-16-hint}
ggplot(data = ..., mapping = aes(x = ..., y = ...)) +
  ...
```

```{r quiz-12, echo = FALSE}
quiz(
  question("Which department has the highest median course rating?",
           answer("COMPSCI"),
           answer("ECON"),
           answer("GOV"),
           answer("MATH", correct = TRUE),
           answer("STAT")
  ),
  question("Which department has outliers?",
           answer("COMPSCI"),
           answer("ECON"),
           answer("GOV"),
           answer("MATH", correct = TRUE),
           answer("STAT")
  ),
  question("Which department has the largest range of course ratings?",
           answer("COMPSCI"),
           answer("ECON"),
           answer("GOV", correct = TRUE),
           answer("MATH"),
           answer("STAT")
  )
)
```

## Barplots

```{r quiz-13, echo = FALSE}
quiz(
  question("A barplot is:",
           answer("A plot that displays the relationship between two numerical variables"),
           answer("A plot that visualizes the distribution of a numerical variable"),
           answer("A plot that visualizes the distribution of a categorical variable",
                  correct = TRUE),
           allow_retry = TRUE)
)
```

###

*We can now use a barplot to plot how many students are enrolled in classes per department. We will use the `top_five_qscores` dataset again to look at the five largest departments.*

```{r barplot-setup, include = FALSE}
top_five_enrolled <- qscores %>% 
  group_by(department) %>%
  summarize(enrollment = sum(enrollment)) %>% 
  arrange(desc(enrollment)) %>%
  pull(department)

top_five_qscores <- qscores %>% filter(department %in% top_five_enrolled[0:5])
```

```{r barplot-top-five-qscores, echo = TRUE}
top_five_qscores
```

*Use `ggplot()` and `geom_bar()` to plot `department` on the x-axis to see the number of classes in each department.*

```{r exercise-17, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-17-hint}
ggplot(data = ..., mapping = aes(x = ...)) +
  ...
```

###

*We can then count the number of classes per department in order to use the `geom_col()` layer instead. We do this using the `group_by()` and `count()` functions in the `tidyverse` package, which you will learn how to use in the next chapter. Our new data frame is stored as `top_five_qscores_counted`.*

```{r classes-counted-setup, include = FALSE}
top_five_qscores_counted <- top_five_qscores %>% 
  group_by(department) %>% count() %>% rename(count = n)
```

```{r avg-workload-setup, echo = TRUE}
top_five_qscores_counted
```

*Use `ggplot()` and `geom_col()` to plot `department` on the x-axis and `count` on the y-axis to see the number of classes in each department.*

```{r exercise-18, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-18-hint}
ggplot(data = top_five_qscores_counted, mapping = aes(x = ..., y = ...)) +
  ...
```

```{r quiz-14, echo = FALSE}
quiz(
  question("What's the difference between `geom_bar()` and `geom_col()`?",
           answer("`geom_bar()` is used to represent pre-counted data while `geom_col() is not."),
           answer("`geom_col()` is used to represent pre-counted data while `geom_bar() is not.", 
                  correct = TRUE),
           answer("There is no difference between the two layers.")
  )
)
```

###

*Let's go back to the `trains` dataset in the `PPBDS.data` package. Use `ggplot()` and `geom_bar()` to plot `treatment` on the x-axis.*

```{r exercise-19, exercise = TRUE, exercise.lines = 5}

```

```{r exercise-19-hint}
ggplot(data = ..., mapping = aes(x = ...)) + ...
```

###

*We can now map the additional variable `party` by adding a `fill = party` inside the `aes()` aesthetic mapping.*

```{r exercise-20, exercise = TRUE, exercise.lines = 5}
ggplot(data = trains, mapping = aes(x = treatment)) + geom_bar()
```

```{r exercise-20-hint}
ggplot(data = trains, mapping = aes(x = treatment, fill = ...)) + geom_bar()
```

###

*Change `fill = party` to `color = party`. What happens? *

```{r exercise-21, exercise = TRUE, exercise.lines = 5}
ggplot(data = trains, mapping = aes(x = treatment, fill = party)) + geom_bar()
```

###

*Create a side-by-side barplot by adding `position = "dodge"` to `geom_bar()`.*

```{r exercise-22, exercise = TRUE, exercise.lines = 5}
ggplot(data = trains, mapping = aes(x = treatment, fill = party)) +
  geom_bar()
```

```{r exercise-22-solution}
ggplot(data = trains, mapping = aes(x = treatment, fill = party)) +
  geom_bar(position = ...)
```

###

*Instead of mapping `party` to `fill`, we can include it as the variable to create small multiples of the plot across the levels of `party` using `facet_wrap(~ party)`. Add this layer to the provided plot in order to create a faceted barplot.*

```{r exercise-23, exercise = TRUE, exercise.lines = 5}
ggplot(data = trains, mapping = aes(x = treatment)) + 
  geom_bar()
```

```{r exercise-23-hint}
ggplot(data = trains, mapping = aes(x = treatment)) + geom_bar() + ...
```

## Submit

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = shiny::div(
    "When you have completed this tutorial, follow these steps:",
        br(),
    tags$ol(
       tags$li("Click Generate Submission"),
       tags$li("Copy and paste the hash generated into the \"Hash\" textbox"),
       tags$li("Press 'Create responses.rds'. Nothing will pop up, but this will create an .rds file of your hashed responses."),
       tags$li("Download the .rds file. A window will pop up asking you where to place the downloaded file. Upload this file to the appropriate Canvas assignment."))
  )
)
```

```{r context="setup"}
fluidPage(
    mainPanel(
          div(id = "form",
            textInput("hash", "Hash", ""),
            actionButton("submit", "Create responses.rds", class = "btn-primary"),
            downloadButton("downloadData", "Download responses.rds")
          )
        )
    )
```

```{r context="server"}

learnrhash::encoder_logic()

 observeEvent(input$submit,{
        hash_id <<- input$hash
        responses <<- data.frame("Hash" = hash_id)
 })
output$downloadData <- downloadHandler(
      filename = "responses.rds",
      content = function(file) {
      write_rds(responses, file)
    }
  )
```


