---
title: "Visualization"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(tidyverse)
library(learnr)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
options(tutorial.storage="client") 
```

## What are R and RStudio?

```{r quiz-1, echo = FALSE}
quiz(
  question("What are R and RStudio?",
             answer("R is a programming language and RStudio is an integrated development environment (IDE)", 
                    correct = TRUE),
             answer("RStudio is a programming language and R is an IDE"),
             answer("R and RStudio are both programming languages"),
             answer("R and RStudio are both IDEs"),
             allow_retry = TRUE
    ),
  question("In what order do you install R and RStudio?",
             answer("First R then RStudio", correct = TRUE),
             answer("First RStudio then R"),
             allow_retry = TRUE
    )
)
```

## How Do I Code in R?

```{r quiz-2, echo = FALSE}
quiz(
  question("Which of the following is NOT the name of a data type in R?",
           answer("Logicals"),
           answer("Integers"),
           answer("Characters"),
           answer("Strings", correct = TRUE),
           allow_retry = TRUE
  ),
  question("What is the operator used in R to test for inclusion?",
           answer("'in'"),
           answer("'%in%'", correct = TRUE),
           answer("'include'"),
           answer("'IN'"),
           allow_retry = TRUE
  ),
  question("What is the symbol used to bring up a help file?",
           answer("*"),
           answer("~"),
           answer("!"),
           answer("?", correct = TRUE),
           allow_retry = TRUE
  ),
  question("True / False: Seeing red text in the console is always bad because it denotes an error.",
           answer("True"),
           answer("False", correct = TRUE, message = "R will show red text in the console pane for errors, warnings, and messages. The only situation in which the code will not run is if there is an error."),
           incorrect = "R will show red text in the console pane for errors, warnings, and messages. The only situation in which the code will not run is if there is an error.",
           allow_retry = TRUE
  )
)
```

## What are R packages?

```{r quiz-3, echo = FALSE}
quiz(
  question("Which of the following is NOT true about R packages?",
           answer("R packages extend the functionality of R by providing additional functions, data, and documentation"),
           answer("If comparing R and R packages to a phone, R would be the apps you can download on the phone and R packages would be the phone", correct = TRUE),
           answer("R packages are written by a worldwide community of R users"),
           answer("R packages can be downloaded for free from the Internet"),
           allow_retry = TRUE
  ),
  question("Which command corresponds to loading a package in R?",
           answer("install.packages(“package”)"),
           answer("library(package)", correct = TRUE),
           allow_retry = TRUE
  )
)
```

*Load the `ggplot2` package.*

```{r exercise-1, exercise = TRUE, exercise.lines = 3}

```

```{r exercise-1-solution}
library(ggplot2)
```

## Explore your first datasets

*Load the `PPBDS.data` package to look at the `trains` dataset.*
```{r exercise-2, exercise = TRUE, exercise.lines = 5}
# ---------
  
trains
```

```{r exercise-2-solution}
library(PPBDS.data)

trains
```

*Use the `glimpse()` function to look at the `trains` dataset. We already loaded the `dplyr` and `PPBDS.data` packages.*
```{r exercise-4, exercise = TRUE, exercise.lines = 5}
library(dplyr)
library(PPBDS.data)

# ---------
```

```{r exercise-4-solution}
library(dplyr)
library(PPBDS.data)

glimpse(trains)
```

```{r quiz-4, echo = FALSE}
quiz(
  question("What is the data type of the `liberal` variable?",
           answer("character"),
           answer("logical", correct = TRUE),
           answer("double"),
           answer("factor"),
           allow_retry = TRUE
  ),
  question("How many rows does the `train` dataset have?",
           answer("8"),
           answer("15"),
           answer("115", correct = TRUE),
           allow_retry = TRUE
  )
)
```

*Extract the `income` variable in the `trains` dataset using the `$` operator.*
```{r exercise-5, exercise = TRUE, exercise.lines = 5}
library(PPBDS.data)

# ---------
```

```{r exercise-5-solution}
library(PPBDS.data)

trains$income
```

*Look at the `qscores` dataset using `glimpse()`. We have loaded the `PPBDS.data` and the `dplyr` packages.*
```{r qscores-setup, include = FALSE}
library(PPBDS.data)
library(dplyr)
```

```{r qscores, echo = TRUE}
glimpse(qscores)
```

```{r quiz-5, echo = FALSE}
quiz(
  question("Which of the following is NOT an identification variable?",
           answer("`course_name`"),
           answer("`department`", correct = TRUE),
           answer("`workload`", correct = TRUE),
           answer("`course_number`"),
           allow_retry = TRUE
  )
)
```

## Components of a graphic

```{r quiz-6, echo = FALSE}
quiz(
  question("What are the three essential components of a graphic?",
           answer("the dataset containing the variables in question", correct = TRUE),
           answer("the geometric object we can observe in a plot", correct = TRUE),
           answer("axes labels on a plot"),
           answer("the aesthetic attributes", correct = TRUE),
           allow_retry = TRUE
  ),
  question("What are the two important arguments that we need to provide the `ggplot()` function?",
           answer("`data` and `mapping`", correct = TRUE),
           answer("`data` and `aesthetics`"),
           answer("`data` and `layers`"),
           allow_retry = TRUE
  )
)
```

*Look at this scatterplot made using the `trains` dataset. We have loaded the `PPBDS.data` and the `ggplot2` packages.*

*Higher numbers mean more conservative attitudes toward immigration-related policies on a 3 to 15 scale. Individuals with a treatment value of "Treated" were exposed to two Spanish-speakers on their regular commute. "Control" individuals were not.*

```{r trains}
library(PPBDS.data)
library(ggplot2)

ggplot(data = trains, mapping = aes(x = att_start, y = att_end, color = treatment)) +
  geom_point() + facet_wrap(~party)
```

```{r quiz-7, echo = FALSE}
quiz(
  question("Which `data` variable is mapped to the `x`-position `aes`thetic of the points?",
           answer("`att_start`", correct = TRUE),
           answer("`att_end`"),
           answer("`treatment`"),
           allow_retry = TRUE
  ),
  question("Which `data` variable is mapped to the `y`-position `aes`thetic of the points?",
           answer("`att_start`"),
           answer("`att_end`", correct = TRUE),
           answer("`treatment`"),
           allow_retry = TRUE
  ),
  question("Which `data` variable is mapped to the `color` `aes`thetic of the points?",
           answer("`att_start`"),
           answer("`att_end`"),
           answer("`treatment`", correct = TRUE),
           allow_retry = TRUE
  )
)
```

## Scatterplots

*Let's make a scatterplot using the `qscores` dataset. Load the `PPBDS.data` package and look at the `qscores` dataset by simply typing the name of the dataset.*

```{r exercise-6, exercise = TRUE, exercise.lines = 5}
# ---------

# ---------
```

```{r exercise-6-solution}
library(PPBDS.data)

qscores
```

*Fill in the `ggplot()` function to create a scatterplot with `overall` on the x-axis and `workload` on the y-axis. Is there a relationship between the overall rating of a Harvard course versus the average hours per week of workload?*

```{r exercise-7, exercise = TRUE, exercise.lines = 5}
library(ggplot2)

# ggplot(data = qscores, mapping = aes(x = ---------, y = ---------)) +
  # geom_point()
```

```{r exercise-7-solution}
ggplot(data = qscores, mapping = aes(x = overall, y = workload)) + 
  geom_point()
```

*Now we want to add a `size` `aes`thetic based on the number of students enrolled in each course. Add `size = enrollment` to the other `aes`thetics of the points. Do larger classes have a higher average overall course rating? What about higher workload?*

```{r exercise-8, exercise = TRUE, exercise.lines = 5}
# ggplot(data = qscores, mapping = aes(x = overall, y = workload, ---------)) +
  # geom_point()
```

```{r exercise-8-solution}
ggplot(data = qscores, mapping = aes(x = overall, y = workload, size = enrollment)) +
  geom_point()
```

*Change the transparency of the points by setting `alpha = 0.2` within `geom_point()`.*

```{r exercise-9, exercise = TRUE, exercise.lines = 5}
# ggplot(data = qscores, mapping = aes(x = overall, y = workload, size = enrollment)) +
  # geom_point(----------)
```

```{r exercise-9-solution}
ggplot(data = qscores, mapping = aes(x = overall, y = workload, size = enrollment)) +
  geom_point(alpha = 0.2)
```

*Replace `geom_point()` with `geom_jitter()` to create a jittered scatterplot.*

```{r exercise-10, exercise = TRUE, exercise.lines = 5}
ggplot(data = qscores, mapping = aes(x = overall, y = workload)) +
  geom_point()
```

```{r exercise-10-solution}
ggplot(data = qscores, mapping = aes(x = overall, y = workload)) +
  geom_jitter()
```

## Linegraphs

```{r linegraphs-setup, include = FALSE}
library(PPBDS.data)
library(tidyverse)

cces_data <- cces %>%
  filter(approval_ch %in% c("Approve / Somewhat Approve", "Disapprove / Somewhat Disapprove")) %>%
  group_by(year, approval_ch) %>%
  summarize(avg_age = mean(age))
```

*The `cces` dataset in the `PPBDS.data` package has been changed to show the average age of voters per year based on their approval of the president that year. The `cces` dataset was manipulated using the `filter()`, `group_by`, and `summarize()` functions within the `tidyverse` package, which you will learn how to use in the next chapter.*

```{r gov_qscores, echo = TRUE}
cces_data
```

*Fill in the `ggplot()` function to create a linegraph with `year` on the x-axis, `avg_age` on the y-axis, and `approval_ch` as the color aesthetic.*

```{r exercise-11, exercise = TRUE, exercise.lines = 5}
library(ggplot2)

# ggplot(data = cces_data, mapping = aes(x = ---------, y = ---------, color = ----------)) +
  # geom_line()
```

```{r exercise-11-solution}
ggplot(data = cces_data, mapping = aes(x = year, y = avg_age, color = approval_ch)) + 
  geom_line()
```

```{r quiz-8, echo = FALSE}
quiz(
  question("True / False: Linegraphs and scatterplots both display the relationship between two numerical variables.",
           answer("True", correct = TRUE),
           answer("False"),
           allow_retry = TRUE),
  question("True / False: It is preferred to use scatterplots over linegraphs when the variable on the x-axis has an inherent ordering, like in a time series plot.",
           answer("True"),
           answer("False", correct = TRUE),
           allow_retry = TRUE)
)
```

## Histograms

```{r quiz-9, echo = FALSE}
quiz(
  question("A histogram is:",
           answer("A plot that displays the relationship between two numerical variables"),
           answer("A plot that visualizes the distribution of a numerical variable", correct = TRUE),
           answer("A plot that visualizes the distribution of a categorical variable"),
           allow_retry = TRUE)
)
```

*Using the `qscores` dataset and `ggplot()`, make a histogram where `overall` is on the x-axis to see the distribution of Harvard course ratings.*

```{r exercise-12, exercise = TRUE, exercise.lines = 5}
library(PPBDS.data)
library(ggplot2)

# ggplot(data = qscores, mapping = aes(x = ---------)) + geom_histogram()
```

```{r exercise-12-solution}
library(PPBDS.data)
library(ggplot2)

ggplot(data = qscores, mapping = aes(x = overall)) + geom_histogram()
```

```{r quiz-10, echo = FALSE}
quiz(
  question_text("What is the default value of bins for this histogram?",
                answer("30", correct = TRUE), 
                allow_retry = TRUE)
)
```

*Add white vertical borders demarcating the bins by adding `color = "powderblue"` and `fill = "violet"` arguments to `geom_histogram()`.*

```{r exercise-13, exercise = TRUE, exercise.lines = 5}
# ggplot(data = qscores, mapping = aes(x = overall)) +
  # geom_histogram(color = --------, fill = --------)
```

```{r exercise-13-solution}
ggplot(data = qscores, mapping = aes(x = overall)) + 
  geom_histogram(color = "powderblue", fill = "violet")
```

*Specify the number of bins to be 10 via the `bins` argument in `geom_histogram()`.*

```{r exercise-14, exercise = TRUE, exercise.lines = 5}
# ggplot(data = qscores, mapping = aes(x = overall)) +
  # geom_histogram(bins = --------, color = "white")
```

```{r exercise-14-solution}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram(bins = 10, color = "white")
```

*Specify the width of the bins to be 1 via the `binwidth` argument in `geom_histogram()`.*

```{r exercise-15, exercise = TRUE, exercise.lines = 5}
# ggplot(data = qscores, mapping = aes(x = overall)) +
  # geom_histogram(binwidth = ---------, color = "white")
```

```{r exercise-15-solution}
ggplot(data = qscores, mapping = aes(x = overall)) +
  geom_histogram(binwidth = 1, color = "white")
```

## Boxplots

```{r quiz-11, echo = FALSE}
quiz(
  question("Which of the following are included in the five-number summary of a numerical variable used to construct a boxplot?",
           answer("Minimum", correct = TRUE),
           answer("First quartile", correct = TRUE),
           answer("Median", correct = TRUE),
           answer("Mean"),
           answer("Third quartile", correct = TRUE),
           answer("Maximum", correct = TRUE)
  )
)
```

*Let's keep playing with the `qscores` dataset in the `PPBDS.data` package. We have filtered the dataset using the `tidyverse` package based on the five departments with the classes with the most enrollment: ECON, COMPSCI, STAT, GOV, and MATH. (You will learn how to do this in the next chapter). The filtered dataset is called `top_five_qscores`.*

```{r boxplot-setup, include = FALSE}
library(tidyverse)
library(PPBDS.data)

top_five_enrolled <- qscores %>% 
  group_by(department) %>%
  summarize(enrollment = sum(enrollment)) %>% 
  arrange(desc(enrollment)) %>%
  pull(department)

top_five_qscores <- qscores %>% filter(department %in% top_five_enrolled[0:5])
```

```{r boxplot-top-five-qscores, echo = TRUE}
top_five_qscores
```

*Using the `top_five_qscores` dataset and `ggplot()`, make a boxplot where `department` is on the x-axis and `overall` is on the y-axis.*

```{r exercise-16, exercise = TRUE, exercise.lines = 5}
library(ggplot2)

# ggplot(data = top_five_qscores, mapping = aes(x = ---------, y = ---------)) + geom_boxplot()
```

```{r exercise-16-solution}
library(ggplot2)

ggplot(data = top_five_qscores, mapping = aes(x = department, y = overall)) + geom_boxplot()
```

```{r quiz-12, echo = FALSE}
quiz(
  question("Which department has the highest median course rating?",
           answer("COMPSCI"),
           answer("ECON"),
           answer("GOV"),
           answer("MATH", correct = TRUE),
           answer("STAT")
  ),
  question("Which department has outliers?",
           answer("COMPSCI"),
           answer("ECON"),
           answer("GOV"),
           answer("MATH", correct = TRUE),
           answer("STAT")
  ),
  question("Which department has the largest range of course ratings?",
           answer("COMPSCI"),
           answer("ECON"),
           answer("GOV", correct = TRUE),
           answer("MATH"),
           answer("STAT")
  )
)
```

## Barplots

```{r quiz-13, echo = FALSE}
quiz(
  question("A barplot is:",
           answer("A plot that displays the relationship between two numerical variables"),
           answer("A plot that visualizes the distribution of a numerical variable"),
           answer("A plot that visualizes the distribution of a categorical variable",
                  correct = TRUE),
           allow_retry = TRUE)
)
```

*We can now use a barplot to plot how many students are enrolled in classes per department. We will use the `top_five_qscores` dataset again to look at the five largest departments.*

```{r barplot-setup, include = FALSE}
library(tidyverse)
library(PPBDS.data)

top_five_enrolled <- qscores %>% 
  group_by(department) %>%
  summarize(enrollment = sum(enrollment)) %>% 
  arrange(desc(enrollment)) %>%
  pull(department)

top_five_qscores <- qscores %>% filter(department %in% top_five_enrolled[0:5])
```

```{r barplot-top-five-qscores, echo = TRUE}
top_five_qscores
```

*Use `geom_bar()` to plot `department` on the x-axis to see the number of classes in each department.*

```{r exercise-17, exercise = TRUE, exercise.lines = 5}
library(ggplot2)

# ggplot(data = top_five_qscores, mapping = aes(x = ---------)) +
  ---------
```

```{r exercise-17-solution}
library(ggplot2)

ggplot(data = top_five_qscores, mapping = aes(x = department)) +
  geom_bar()
```

*We can then count the number of classes per department in order to use the `geom_col()` layer instead. We do this using the `group_by()` and `count()` functions in the `tidyverse` package, which you will learn how to use in the next chapter. Our new data frame is stored as `top_five_qscores_counted`.*

```{r classes-counted-setup, include = FALSE}
library(tidyverse)

top_five_qscores_counted <- top_five_qscores %>% group_by(department) %>% count() %>% rename(count = n)
```

```{r avg-workload-setup, echo = TRUE}
top_five_qscores_counted
```

*Use `geom_col()` to plot `department` on the x-axis and `count` on the y-axis to see the number of classes in each department.*

```{r exercise-18, exercise = TRUE, exercise.lines = 5}
# ggplot(data = top_five_qscores_counted, mapping = aes(x = ---------, y = ---------)) +
  # ---------
```

```{r exercise-18-solution}
ggplot(data = top_five_qscores_counted, mapping = aes(x = department, y = count)) +
  geom_col()
```

```{r quiz-14, echo = FALSE}
quiz(
  question("What's the difference between `geom_bar()` and `geom_col()`?",
           answer("`geom_bar()` is used to represent pre-counted data while `geom_col() is not."),
           answer("`geom_col()` is used to represent pre-counted data while `geom_bar() is not.", 
                  correct = TRUE),
           answer("There is no difference between the two layers.")
  )
)
```

*Let's go back to the `trains` dataset in the `PPBDS.data` package. Use `geom_bar()` to plot `treatment` on the x-axis.*

```{r exercise-19, exercise = TRUE, exercise.lines = 5}
library(PPBDS.data)
library(ggplot2)
 
# ggplot(data = trains, mapping = aes(x = ---------)) + ---------
```

```{r exercise-19-solution}
library(PPBDS.data)
library(ggplot2)
 
ggplot(data = trains, mapping = aes(x = treatment)) + geom_bar()
```

*We can now map the additional variable `party` by adding a `fill = party` inside the `aes()` aesthetic mapping.*

```{r exercise-20, exercise = TRUE, exercise.lines = 5}
# ggplot(data = trains, mapping = aes(x = treatment, fill = --------)) + geom_bar()
```

```{r exercise-20-solution}
ggplot(data = trains, mapping = aes(x = treatment, fill = party)) + geom_bar()
```

*Change `fill = party` to `color = party`. What happens? *

```{r exercise-21, exercise = TRUE, exercise.lines = 5}
ggplot(data = trains, mapping = aes(x = treatment, fill = party)) + geom_bar()
```

```{r exercise-21-solution}
ggplot(data = trains, mapping = aes(x = treatment, color = party)) + geom_bar()
```

*Create a side-by-side barplot by adding `position = "dodge"` to `geom_bar()`.*

```{r exercise-22, exercise = TRUE, exercise.lines = 5}
# ggplot(data = trains, mapping = aes(x = treatment, fill = party)) + geom_bar(---------)
```

```{r exercise-22-solution}
ggplot(data = trains, mapping = aes(x = treatment, fill = party)) + geom_bar(position = "dodge")
```

*Instead of mapping `party` to `fill`, we can include it as the variable to create small multiples of the plot across the levels of `party` using `facet_wrap(~ party)`. Add this layer to the provided plot in order to create a faceted barplot.*

```{r exercise-23, exercise = TRUE, exercise.lines = 5}
# ggplot(data = trains, mapping = aes(x = treatment)) + geom_bar() + ---------
```

```{r exercise-23-solution}
ggplot(data = trains, mapping = aes(x = treatment)) + geom_bar() + facet_wrap(~ party)
```

<!-- ## Adding Layers to a Plot -->

<!-- Need to add: faceting, position, scales, coordinate systems, statistical transformations, adding titles and labels to a graph -->


