---
title: "Probability"
tutorial:
  id: "05-probability"
output:
  learnr::tutorial:
      progressive: true
      allow_skip: true
runtime: shiny_prerendered
description: "Chapter 5 tutorial"
---

```{r setup, include=FALSE}
library(tidyverse)
library(PPBDS.data)
library(learnr)
library(shiny)
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
options(tutorial.storage="local")  
```

## Name

```{r name, echo=FALSE}
question_text(
  "Student Name:",
  answer(NULL, correct = TRUE),
  incorrect = "Ok",
  try_again_button = "Modify your answer",
  allow_retry = TRUE
)
```




## Flipping a coin
###

Before we begin modeling a coin flip, let's make sure we understand the important vocabulary.

```{r coin1}
question_text("In a sentence, explain the difference between an empirical distrubtion and a posterior distribution.",
              answer(NULL, correct = TRUE), 
              incorrect = "Submitted.",
              message = "A posterior distribution is based on what you believe the distribution of some data will be, whereas an empiral distribution is the distribution of some actual data.")
```

###

Suppose you have designed an experiment to gather data on a topic.

```{r coin2}
question_text("How many times must you conduct the experiment to create a mathemetical distribution?",
              answer(NULL, correct = TRUE), 
              incorrect = "Submitted.",
              message = "A mathematical distrubition requires no experimental data because it relies solely on mathematical formulas. Thus, you don't need to conduct the experiment at all before creating a mathetmatical distribution.")
```

###

The simplest way to visualize probabilities is with a coin flip.

Suppose for some coin that $p_t = 0.4$.

```{r coin3}
question_text("What does the above statement imply about the coin?",
              answer(NULL, correct = TRUE), 
              incorrect = "Submitted.",
              message = "Because the probability of tails is 0.4 (or 40%), the coin is more likely to come up heads than tails. Therefore, the coin must not be a fair coin.")
```

###

Let's start by examining the standard, unweighted coin: i.e. $p_h = p_t = 0.5$. 

The function `rbinom()` is used to calculate a binomial distribution: i.e. when there are only two possible outcomes. Look at the following code to see how we can simulate a single coin flip. Try running the code multiple times---see if you can estimate approximately how many times you get each outcome.

```{r coin4, exercise = TRUE}
rbinom(n = 1, size = 1, prob = 0.5)
```

###

In the code chunk below, change our default `rbinom()` statement such that the number of *observations* (the `n` parameter) is equal to 100.

```{r coin5, exercise = TRUE}
rbinom(n = 1, size = 1, prob = 0.5)
```

###

This time, change the number of *trials* (the `size` parameter) to 100. See if you can figure out the difference between observations and trials in the `rbinom()` function.

```{r coin5_2, exercise = TRUE}
rbinom(n = 1, size = 1, prob = 0.5)
```

###

Write an `rbinom()` statement to simulate the flipping of a coin that lands on heads 25% of the time (i.e. set `prob` to 0.25). Your simulation should have 50 observations and 10 trials per observation.

```{r coin5_3, exercise = TRUE}

```

```{r coin5_3-hint, eval=FALSE}
Remember that n is the number of observations and size is the number of trials.
```

###

Use `tibble()` to create a new tibble. The tibble should have one column, `head`, equal to a binomial distribution with 100 observations of a fair coin (`prob = 0.5`) and 1 trial (i.e. 1 flip) per observation.

```{r coin6, exercise = TRUE}

```

```{r coin6-hint, eval=FALSE}
tibble(head = ...)
```

###

Pipe your tibble into `ggplot()` to make a histogram with `head` mapped to the x-axis. Remember to copy and paste your code from the previous exercise.

```{r coin7, exercise = TRUE}

```

###

To improve the plot's aesthetics, set the `binwidth` argument of `geom_histogram()` to 0.5.

```{r coin8, exercise = TRUE, exercise.lines = 4}

```

###

We can use a `scale_x_` function, along with the `breaks` and `labels` parameters, to reset the tick marks on the x-axis. Run the following code to see how to reset our graph's tick mark labels at x = 0 and x = 0.5.

```{r coin9, exercise = TRUE}
tibble(head = rbinom(n = 100, size = 1, prob = 0.5)) %>%
  ggplot(mapping = aes(x = head)) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = c(0, 0.5), labels = c("Heads", "Random Label With No Bar"))
```


###

Use `scale_x_continuous()` like we did in the previous code chunk to reset the tick mark labels at x = 0 and x = 1 to "Heads" and "Tails." In addition, change the theme of the plot to `theme_classic()`.

```{r coin10, exercise = TRUE, exercise.lines = 4}
tibble(head = rbinom(n = 100, size = 1, prob = 0.5)) %>%
  ggplot(mapping = aes(x = head)) +
    geom_histogram(binwidth = 0.5)
```

```{r coin10-hint, eval = FALSE}
tibble(head = rbinom(n = 100, size = 1, prob = 0.5)) %>%
  ggplot(mapping = aes(x = head)) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = ..., labels = ...) +
    ...
```

###

Finally, use `labs()` to set the following labels: the graph's `title` to "Empirical Probability Distribution," the graph's `subtitle` to "Flipping A Coin One Hundred Times," and the graph's `x`-axis label to "Outcome."

You have now made a very good copy of the graph in the textbook---great work!

```{r coin11, exercise = TRUE, exercise.lines = 6}

```

```{r coin11-hint, eval = FALSE}
tibble(head = rbinom(n = 100, size = 1, prob = 0.5)) %>%
  ggplot(mapping = aes(x = head)) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = c(0, 1), labels = c("Heads", "Tails")) +
    theme_classic() +
    labs(...)
```

###

Let's imagine we ran the code above 1,000 times. Most of the time, we'd probably get a result fairly close to 50 heads and 50 tails, while sometimes (just from random chance), there would be significantly more heads than tails, or vice versa. But exactly how often would each unique distribution occur?

We're now going to make a different kind of empirical distribution. Start by using `tibble()` again to make a tibble. The sole column of the tibble should be `heads`, equal to the result of an `rbinom()` statement with 1,000 observations of a fair coin and 100 trials (flips) per observation.

```{r coin12, exercise = TRUE}

```

```{r coin12-hint-1, eval=FALSE}
To simulate a fair coin, set the prob argument of rbinom() to 0.5.
```

```{r coin12-hint-2, eval = FALSE}
tibble(heads = ...)
```

###

Pipe your tibble into `ggplot()` to make a histogram with `heads` mapped to the x-axis.

```{r coin13, exercise = TRUE}

```

###

For a clearer view of the distribution, set the `binwidth` of the histogram to 1.

```{r coin14, exercise = TRUE, exercise.lines = 4}

```

###

Great work! This empirical distribution shows that while there is a significant amount of variance in the number of heads in 100 coin flips, nearly all of the observations have greater than 40 heads but less than 60 heads.

For one last visual, weight the coin by changing the `prob`ability of heads to 0.95 and observe the resulting distribution.

```{r coin15, exercise = TRUE, exercise.lines = 4}
tibble(heads = rbinom(n = 1000, size = 100, prob = 0.5)) %>%
  ggplot(mapping = aes(x = heads)) +
    geom_histogram(binwidth = 1)
```

```{r coin15-hint}
# Change the prob argument of rbinom() from 0.5 to 0.95.
```


## Rolling two dice
###

For a step up in complexity, let's look at the probability distributions of one or more dice rolls. This is more difficult to model because we cannot simply use `rbinom()` to make a binomial distribution (there are six outcomes, not two).

Recall from chapter 2 that we use the `sample()` function to generate random values. Run the code chunk below to see how we can generate ten random coin tosses using `sample()`. Remember to set `replace` equal to `TRUE` when you want a certain outcome (i.e. "heads" or "tails" in this case) to be achievable more than once.

```{r dice1, exercise = TRUE}
sample(x = c("heads", "tails"), size = 10, replace = TRUE)
```

###

Using `sample()`, simulate 10 dice rolls by setting the `x` argument to the vector of integers from 1 through 6. Note that you don't actually need to use the `c()` function in this case.

```{r dice2, exercise = TRUE}

```

```{r dice2-hint-1}
# Use the same code as in the previous exercise, but change the value of sample()'s x parameter.
```

```{r dice2-hint-2}
# 1:6 is equivalent to c(1, 2, 3, 4, 5, 6).
```

###

Next, to make 10 simulations of the sum of *two* dice rolls, simply add together two of the `sample()` statements that you made in the last exercise.

```{r dice3, exercise = TRUE}

```

```{r dice3-hint}
# Just add the sample statement from the previous exercise to itself.
```

###

Use `tibble()` to create a new tibble with one variable, `sum`, equal to the sum of the two `sample()` statements that you made in the last exercise. Remember to copy and paste your code from the previous exercise!

```{r dice4, exercise = TRUE}

```

```{r dice4-hint, eval = FALSE}
tibble(sum = ...)
```

###

Add a `ggplot()` call, using a pipe, to create a histogram with `sum` mapped to the x-axis.

```{r dice5, exercise = TRUE}

```

###

This graph is really difficult to read. To improve it, change `geom_histogram()`'s `binwidth` argument to 1 and its `color` argument to "white."

```{r dice6, exercise = TRUE, exercise.lines = 5}

```

###

Just as we did in the coin toss section, use `scale_x_continuous` to reset the tick mark labels on the y-axis. Both the `breaks` and `labels` parameters should be equal to the vector containing all of the integers from 2 to 12 (inclusive).

```{r dice7, exercise = TRUE, exercise.lines = 5}

```

```{r dice7-hint, eval = FALSE}
tibble(sum = sample(x = 1:6, size = 10, replace = TRUE) +
                           sample(x = 1:6, size = 10, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = ..., labels = ...)
```

###

Finally, finish your graph by changing its theme to `theme_classic()`. In addition, use `labs()` to change the `title` of the graph to "Empirical Probability Distribution," the `subtitle` to "Sum from Rolling Two Dice Ten Times," and the `x`-axis label to "Sum of Two Dice."

```{r dice8, exercise = TRUE, exercise.lines = 6}

```

```{r dice8-hint, eval = FALSE}
tibble(sum = sample(x = 1:6, size = 10, replace = TRUE) +
                           sample(x = 1:6, size = 10, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    ... +
    labs(...)
```

###

Great work! Notice that while sums around 7 are the most common, there is an incredible amount of variance with only 10 rolls.

To see how variance decreases with an increased sample size, copy and paste your completed graph to make a second graph below the first one. The second graph should be identical to the first one, except for two things: the `size` argument in **both** of the `sample()` statements of the second graph should be equal to 100 instead of 10, and the subtitle of the second graph should replace "Ten Times" with "One Hundred Times."

```{r dice9, exercise = TRUE, exercise.lines = 10}
tibble(sum = sample(x = 1:6, size = 10, replace = TRUE) +
                           sample(x = 1:6, size = 10, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    theme_classic() +
    labs(title = "Empirical Probability Distribution",
         subtitle = "Sum from Rolling Two Dice Ten Times",
         x = "Sum of Two Dice")
```

```{r dice9-hint, eval = FALSE}
tibble(sum = sample(x = 1:6, size = 10, replace = TRUE) +
                           sample(x = 1:6, size = 10, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
  geom_histogram(binwidth = 1, color = "white") +
  scale_x_continuous(breaks = 2:12, labels = 2:12) +
  theme_classic() +
  labs(title = "Empirical Probability Distribution",
       subtitle = "Sum from Rolling Two Dice Ten Times",
       x = "Sum of Two Dice")

tibble(sum = sample(x = 1:6, size = ..., replace = TRUE) +
                           sample(x = 1:6, size = ..., replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    theme_classic() +
    labs(title = "Empirical Probability Distribution",
         subtitle = ...,
         x = "Sum of Two Dice")
```

###

The code chunk below contains only the second graph from the previous exercise (the graph with 100 rolls). I have also changed the first `sample()` statement such that the first die has a 50% chance of rolling a 1 and a 10% chance of rolling each other number.

To finish this section, change the `prob` argument of the second `sample()` statement such that there is a 50% chance of rolling a 6 and a 10% chance of rolling each other number. Try to predict what the resulting distribution will look like before you run the code.

```{r dice10, exercise = TRUE, exercise.lines = 10}
tibble(sum = sample(x = 1:6, size = 100, replace = TRUE, prob = c(0.5, 0.1, 0.1, 0.1, 0.1, 0.1)) +
                           sample(x = 1:6, size = 100, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    theme_classic() +
    labs(title = "Empirical Probability Distribution",
         subtitle = "Sum from Rolling Two Dice One Hundred Times",
         x = "Sum of Two Dice")
```

```{r dice10-hint, eval = FALSE}
tibble(sum = sample(x = 1:6, size = 100, replace = TRUE, prob = c(0.5, 0.1, 0.1, 0.1, 0.1, 0.1)) +
                           sample(x = 1:6, size = 100, replace = TRUE, prob = ...)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    theme_classic() +
    labs(title = "Empirical Probability Distribution",
         subtitle = "Sum from Rolling Two Dice One Hundred Times",
         x = "Sum of Two Dice")
```











## Normalized distributions

If you've payed close attention to this tutorial, you might notice that the y-axes of the graphs are slightly different than the y-axes of the graphs in the textbook. This is because the graphs in the textbook are *normalized*.

```{r normalize1}
question_text("In a sentence, explain the difference between normalized and unnormalized distributions.",
              answer(NULL, correct = TRUE), 
              incorrect = "Submitted.",
              message = "A normalized distribution will display the probability of each outcome on the y-axis, whereas an unnormalized distribution will display the count of the number of times that outcome occurs.")
```

###

To make a normalized distribution, we'll need to manipulate the `count` variable of our histograms. Unfortunately, the `count` variable is created by ggplot2---it does not actually exist in our data set!

To see why this is a problem, try running the following code, which attempts to make a histogram of dice rolls and then see how then see how the count for each roll differs from the average count. Then, continue onto the next exercise to see how we can fix the error.

```{r normalize2, exercise = TRUE, eval = FALSE, exercise.lines = 4}
tibble(roll = sample(x = 1:6, size = 20, replace = TRUE)) %>%
  ggplot(mapping = aes(x = roll, y = count / mean(count))) +
    geom_histogram(binwidth = 1, color = "white")
```

###

To tell R that `count` is not actually in our data set but is created by `geom_histogram()`, we need to surround the variable by two dots on each side (`..count..`). Use this technique to adjust the code below so that there is no error.

```{r normalize3, exercise = TRUE, eval = FALSE, exercise.lines = 4}
tibble(roll = sample(x = 1:6, size = 20, replace = TRUE)) %>%
  ggplot(mapping = aes(x = roll, y = count / mean(count))) +
    geom_histogram(binwidth = 1, color = "white")
```


###

Finally, to make a normalized distribution, copy and paste your code from the previous exercise, but divide each `count` by the `sum()` of the counts instead of the `mean()` of the counts. This is the very definition of probability: the number of outcomes of an event divided by the total number of outcomes of all possible events.

```{r normalize4, exercise = TRUE, exercise.lines = 4}

```

```{r normalize4-hint, eval = FALSE}
dice_roll <- tibble(roll = sample(x = 1:6, size = 20, replace = TRUE)) %>%
  ggplot(mapping = aes(x = roll, y = ...)) +
    geom_histogram(binwidth = 1, color = "white")
```

###

The code below makes the graph of the sum from rolling two dice 100 times that we made in the previous section. Notice how the graph currently shows counts, not probabilities, on the y-axis.

Use the technique from the previous exercise to normalize the distribution.

```{r normalize5, exercise = TRUE, exercise.lines = 10}
tibble(sum = sample(x = 1:6, size = 100, replace = TRUE) +
                           sample(x = 1:6, size = 100, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    theme_classic() +
    labs(title = "Empirical Probability Distribution",
         subtitle = "Sum from Rolling Two Dice One Hundred Times",
         x = "Sum of Two Dice")
```

```{r normalize5-hint, eval = FALSE}
tibble(sum = sample(x = 1:6, size = 100, replace = TRUE) +
                           sample(x = 1:6, size = 100, replace = TRUE)) %>%
  ggplot(mapping = aes(x = sum, y = ...)) +
    geom_histogram(binwidth = 1, color = "white") +
    scale_x_continuous(breaks = 2:12, labels = 2:12) +
    theme_classic() +
    labs(title = "Empirical Probability Distribution",
         subtitle = "Sum from Rolling Two Dice One Hundred Times",
         x = "Sum of Two Dice")
```

###

Great work! You now know how to make a normalized distribution. To finish the plot from the textbook, copy and paste your code from the previous exercise, but change the y-axis label to "Probability" using the `y` parameter of `labs()`.

```{r normalize6, exercise = TRUE, exercise.lines = 10}
two_dice_rolls <- tibble(sum = sample(x = 1:6, size = 100, replace = TRUE) +
                           sample(x = 1:6, size = 100, replace = TRUE))
ggplot(two_dice_rolls, mapping = aes(x = sum, y = ..count.. / sum(..count..))) +
  geom_histogram(binwidth = 1, color = "white") +
  scale_x_continuous(breaks = 2:12, labels = 2:12) +
  theme_classic() +
  labs(title = "Empirical Probability Distribution",
       subtitle = "Sum from Rolling Two Dice One Hundred Times",
       x = "Sum of Two Dice")
```

```{r normalize6-hint, eval = FALSE}
two_dice_rolls <- tibble(sum = sample(x = 1:6, size = 100, replace = TRUE) +
                           sample(x = 1:6, size = 100, replace = TRUE))
ggplot(two_dice_rolls, mapping = aes(x = sum, y = ..count.. / sum(..count..))) +
  geom_histogram(binwidth = 1, color = "white") +
  scale_x_continuous(breaks = 2:12, labels = 2:12) +
  theme_classic() +
  labs(title = "Empirical Probability Distribution",
       subtitle = "Sum from Rolling Two Dice One Hundred Times",
       x = "Sum of Two Dice",
       ...)
```










## Two models
###

The classic example of two models is for disease testing. Suppose that 1% of a city with population 10,000 has a deadly disease. The test for the disease is fairly accurate: if you have the disease, the test will correctly diagnose you 99% of the time. If you don't have the disease, the test will correctly diagnose you 95% of the time. The question we want to answer is: what is the probability that you have the disease, given that you test positive?

To start, simply call an `rbinom()` statement with 10,000 observations, 1 trial per observation, and `prob` equal to 0.01. This will make a vector of length 10,000 such that every person infected by the disease (probability 0.01) is represented by the number 1, and every person not infected is represented by the number 0.

```{r twomodels1, exercise = TRUE}

```

```{r twomodels1-hint-1}
# Remember that n is the number of observations and size is the number of trials per observation.
```

```{r twomodels1-hint-2, eval = FALSE}
rbinom(n = ..., size = ..., prob = ...)
```

###

Using `tibble()`, create a tibble with one column, `has_disease`. `has_disease` should be equal to the result of the `rbinom()` statement from the previous exercise.

```{r twomodels2, exercise = TRUE}

```

```{r twomodels2-hint, eval = FALSE}
tibble(...)
```

###

Start a pipe and `mutate()` a new variable, `test_positive`. `test_positive` should be equal the result of an `if_else()` statement with the condition being that `has_disease` is equal to 1. For now, if the condition is true, set `test_positive` to 1, but if the condition is false, set `test_positive` to 0 (we'll put in the exact probabilities in the next exercise).

```{r twomodels3, exercise = TRUE}

```

```{r twomodels3-hint, eval = FALSE}
tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(...,
                                 true = 1,
                                 false = 0))
```

###

According to our current model, every person who has the disease automatically tests positive, and vice versa for those who don't. To add the possibility of a false positive or false negative, we need to add two more `rbinom()` statements inside our `if_else()` statement.

Use the model I gave you in the `true` case to add an `rbinom()` statement to the `false` case. The `rbinom()` statement should have 10,000 observations, 1 trial per observation, and a probability of testing positive equal to 0.05 (since the test correctly diagnoses a negative result for an uninfected person 95% of the time).

```{r twomodels4, exercise = TRUE, exercise.lines = 5}
tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = 0L))
```

```{r twomodels4-hint, eval = FALSE}
tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(...)))
```

###

Note the use of 0L instead of 0. This is a sutble issue. `if_else()` is a safer (i.e., more strict) implementation of `ifelse()`. `if_else()` requires that the type of the variable always be the same. `ifelse()`, on the other hand, allows whatever types you like, and then coerces in a sensible way. `rbinom()` returns an integer for the `true` argument. So, `if_else()` requires `false` to also be an integer. But a simple 0 is read as a double. In order to force R to treat it as an integer, we need the `L` at the end.

Awesome work! We now have a tibble that uses randomness to create a model of our disease.

Use the assignment operator (`<-`) to save the tibble from the previous exercise to `disease_testing`.

```{r twomodels5, exercise = TRUE, exercise.lines = 5}

```

###

Add a `ggplot()` call to make a jittered scatterplot that takes in `disease_testing` as its data, maps `test_positive` to the x-axis, and maps `has_disease` to the y-axis.

```{r twomodels6, exercise = TRUE, exercise.lines = 5}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))
```

```{r twomodels6-hint, eval = FALSE}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))

ggplot(...) +
  ...
```

###

Change the theme of the graph to `theme_classic()`. In addition, use `labs()` to set the graph's `title` to "Joint Empirical Distribution of Test Result and Disease Status," `x`-axis label to "Test Result," and `y`-axis label to "Disease Status."

```{r twomodels7, exercise = TRUE, exercise.lines = 8}

```

```{r twomodels7-hint, eval = FALSE}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))

ggplot(disease_testing, mapping = aes(x = test_positive, y = has_disease)) +
  geom_jitter() +
  ... +
  labs(...)
```

###

Finally, use `scale_x_continuous()` and `scale_y_continuous()` to reset the tick mark labels on both axes. The `breaks` for both axes should be 0 and 1, and the `labels` for both axes should be "Negative" and "Positive."

```{r twomodels8, exercise = TRUE, exercise.lines = 12}

```

```{r twomodels8-hint-1}
# In both scale_ calls, the breaks argument should be a numeric vector and the labels argument should be a character vector.
```

```{r twomodels8-hint-2, eval = FALSE}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))

ggplot(disease_testing, mapping = aes(x = test_positive, y = has_disease)) +
  geom_jitter() +
  theme_classic() +
  labs(title = "Joint Empirical Distribution of Test Result and Disease Status",
       x = "Test Result",
       y = "Disease Status") +
  scale_x_continuous(...) +
  scale_y_continuous(...)
```

###

Fantastic job! You have now made a joint empirical distribution just like the graph in the textbook. This graph makes it pretty clear that you are more likely *not* to have the disease even if you test positive. However, it's hard to tell exactly *how* much likelier. For that, we're going to need a histogram.

Start by making a new pipe with `disease_testing` and `filter()` the pipe to just the rows such that `test_positive` is equal to 1.

```{r twomodels9, exercise = TRUE, exercise.lines = 7}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))

disease_testing
```

```{r twomodels9-hint, eval = FALSE}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))

disease_testing %>%
  ...
```

###

Add a `ggplot()` call to the pipe to make a histogram that maps the `has_disease` variable to the x-axis.

```{r twomodels10, exercise = TRUE, exercise.lines = 8}

```

###

Improve the aesthetics of the plot by setting the `binwidth` of the histogram to 0.5.

```{r twomodels11, exercise = TRUE, exercise.lines = 10}

```

###

Use `scale_x_continuous()` just like we did in the joint distribution graph to reset the tick mark labels on the x-axis. Again, we want the `breaks` to be 0 and 1 and the `labels` to be "Negative" and "Positive."

```{r twomodels12, exercise = TRUE, exercise.lines = 10}

```

###

Next, use the same two-dot technique that we used in the previous section to normalize the distribution (we want to divide the `count` of each bar by the `sum()` of the counts of all the bars).

```{r twomodels13, exercise = TRUE, exercise.lines = 11}

```

```{r twomodels13-hint-1}
# Remember to replace "count" with "..count.." whenever you use it.
```

```{r twomodels13-hint-2, eval = FALSE}
disease_testing %>%
  filter(test_positive == 1) %>%
  ggplot(mapping = aes(x = has_disease, y = ...)) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = 0:1, labels = c("Negative", "Positive"))
```

###

Finally, use `labs()` to set the `title` of the graph to "Conditional Probability Distribution," the `subtitle` to "Probability of Having the Disease Given a Positive Test Result," the `x`-axis label to "Disease Status," and the `y`-axis label to "Probability."

```{r twomodels14, exercise = TRUE, exercise.lines = 11}

```

```{r twomodels14-hint, eval = FALSE}
disease_testing %>%
  filter(test_positive == 1) %>%
  ggplot(mapping = aes(x = has_disease, y = ..count.. / sum(..count..))) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = 0:1, labels = c("Negative", "Positive")) +
    labs(...)
```

###

Great work! From this graph, we can see that you have between an 80% and 85% chance of *not* having the disease, even after you test positive.

To finish the graph as shown in the textbook, change the theme to `theme_classic()`.

```{r twomodels15, exercise = TRUE, exercise.lines = 15}
disease_testing <- tibble(has_disease = rbinom(n = 10000, size = 1, prob = 0.01)) %>%
  mutate(test_positive = if_else(condition = has_disease == 1,
                                 true = rbinom(n = 10000, size = 1, prob = 0.99),
                                 false = rbinom(n = 10000, size = 1, prob = 0.05)))

disease_testing %>%
  filter(test_positive == 1) %>%
  ggplot(mapping = aes(x = has_disease, y = ..count.. / sum(..count..))) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = 0:1, labels = c("Negative", "Positive")) +
    labs(title = "Conditional Probability Distribution",
         subtitle = "Probability of Having the Disease Given a Positive Test Result",
         x = "Disease Status",
         y = "Probability")
```

```{r twomodels15-hint, eval = FALSE}
disease_testing %>%
  filter(test_positive == 1) %>%
  ggplot(mapping = aes(x = has_disease, y = ..count.. / sum(..count..))) +
    geom_histogram(binwidth = 0.5) +
    scale_x_continuous(breaks = 0:1, labels = c("Negative", "Positive")) +
    labs(title = "Conditional Probability Distribution",
         subtitle = "Probability of Having the Disease Given a Positive Test Result",
         x = "Disease Status",
         y = "Probability") +
    ...
```



## Submit

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
