---
title: "Shopping Week"
output:
  learnr::tutorial:
      progressive: true
      allow_skip: true
runtime: shiny_prerendered
description: "Complete a simple tutorial during the first class to ensure your computer is set up correctly."
---

```{r setup, include=FALSE}
library(tidyverse)
library(learnr)
library(learnrhash)
library(shiny)
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
options(tutorial.exercise.timelimit = 60)  
```

## Chapter 00 tutorial

<!-- MB: How to make question not say "incorrect" to names -->

Welcome to your first tutorial. The goal of this exercise is to answer some questions about the shopping week chapter and to learn how to complete and submit a tutorial. Begin by filling out your name and clicking the submit button. If you made a mistake, you click "Try Again" and edit your name. 
``` {r name, echo=FALSE}
question_text(
  "Student Name:",
  answer(NULL, correct = TRUE),
  allow_retry = TRUE
)
```

### Exercise 1
For our first exercise, we'll do some simple math in our console. We've already assigned variables x to 5 and y to 4. On the next line, add the two variables. Hit "Run Code" to see the output of your addition. Clicking "Hint" gives you a hint about what to type into the console. Clicking "Start Over" clears the console and allows you to begin again.

```{r question-1-1, exercise = TRUE}
x <- 5
y <- 4
```

```{r question-1-1-hint, silent = TRUE}
x + y
```


### Exercise 2
Check the version of R running on your computer. You can do this by copy-pasting `version[['version.string']]` into your console and hit "Run Code". You will use the output of this line of code to answer the following question. 
```{r question-1-2, exercise = TRUE}

```

```{r question-1-2-hint, silent = TRUE}
version[['version.string']]
```

Here is a different type of question you will encounter in tutorials: multiple choice. If you get the wrong answer for this question, you haven't installed the most recent version of R and should redo the steps from chapter 00.
If you click the wrong answer in general on multiple choice questions, you can hit "Try Again" for another attempt. 
```{r question-1-3}

question("What is the R version running on your computer?",
    answer("4.0.2"),
    answer("4.0.1", correct = TRUE),
    answer("4.0.0"),
    answer("3.9.9"),
    allow_retry = TRUE
  )

```

### Exercise 3

At this point, we want to know if your machine saves your progress on tutorials. Close your tutorial and restart your R session, then open up this tutorial again. Was your progress on exercises 1 and 2 saved? If not, ask for help on setting up tutorials.

### Exercise 4
Now, we are going to load our first dataset. We've already loaded the PPBDS.data in the package -- this is the package our dataset will come from. Type `trains` in the console and hit "Run Code". This should display some observations from the trains dataset. 

```{r question-1-4, exercise = TRUE}
library(PPBDS.data)
```

```{r question-1-4-hint, silent = TRUE}
trains
```

<!-- Get rid of this exercise? -->

### Exercise 5
Check if your computer is set to download R packages in binary by typing in `options("pkgType")`. Does it say binary?
```{r question-1-5, exercise = TRUE}

```

```{r question-1-5-hint, silent = TRUE}
options("pkgType")
```

## Submit

```{r encode, echo=FALSE}
learnrhash::encoder_ui(
  ui_before = shiny::div(
    "When you have completed this tutorial, follow these steps:",
        br(),
    tags$ol(
       tags$li("Click Generate Submission"),
       tags$li("Copy and paste the hash generated into the \"Hash\" textbox"),
       tags$li("Press 'Create responses.rds'. Nothing will pop up, but this will create an .rds file of your hashed responses."),
       tags$li("Download the .rds file. A window will pop up asking you where to place the downloaded file. Upload this file to the appropriate Canvas assignment."))
  )
)
```

```{r context="setup"}
fluidPage(
    mainPanel(
          div(id = "form",
            textInput("hash", "Hash", ""),
            actionButton("submit", "Create responses.rds", class = "btn-primary"),
            downloadButton("downloadData", "Download responses.rds")
          )
        )
    )
```

```{r context="server"}

learnrhash::encoder_logic()

 observeEvent(input$submit,{
        hash_id <<- input$hash
        responses <<- data.frame("Hash" = hash_id)
 })
output$downloadData <- downloadHandler(
      filename = "tutorial_0_responses.rds",
      content = function(file) {
      write_rds(responses, file)
    }
  )
```

